% !TEX root = main.tex
% Notes for Prof. Santiago:
% I am trying to figure out a way to neatly section everything off along with adding
% some sort of standardized academic paper formatting to the document.
% I am also debating a single or double column layout for the report. 
% 

\documentclass[12pt, letterpaper]{article}
\usepackage[T1]{fontenc}

\title{Exploring the use of differential equations in electric circuits}

\begin{document}

Abstract

In this paper, we explore series RLC circuits and the differential equations that power them. 
We analyze the use of first and second order differential equations in modeling series RLC equations, 
along with exploring real-world use cases and connecting them to these differential equations. 

Key Words: 

- Series RLC circuit
- Differential equations


Table of Contents

1) Introduction

2) Circuit equations

In electric circuit theory, a series RLC circuit is a closed loop circuit consisting of a resistor, inductor and capacitor in series along with a voltage source. 

The source material being used to primarily derive equations for the series RLC circuit are provided from Nilsson and Riedel's Electric Circuits 12th edition textbook.
This textbook is primarily assigned as the main reference and material for ENGR 230 and CMPE 330 at Cal State East Bay, and relies on linear algebra and differential equations as 
co-requisite courses. 

2.1 Circuit Overview 

We consider a series RLC circuit driven by a voltage source V(t). The quantity of interest is 
the capacitor voltage vCt for t>= 0.

We will assume the given component values:

Source voltage: 20V

Resistance: 40 Ohms $ \Omega $

Inductance: 100 mH

Capacitance: 25 $ \mu $ F

Now that we have these variables in place, let's set up the tools we will need for our diff eq.

State variables

The two state variables here are the voltage across the capacitor and the current through the inductor, which also represents the current across the circuit.

We choose these two variables as the capcaitor's voltage and inductor's current cannot change instantaneously, which makes them suitable for representing the system's
energy state at any instant in time. 

Initial conditions

Our circuit will initially have a fully charged capacitor at 20 volts, and the inductor's current will start at zero. 

Before the switch is flipped, the capacitor will behave like an open circuit and the inductor behaves like a short circuit. 

Governing equations

Kirchhoff's Voltage Law 

Deriving Kirchhoff's Voltage Law from the textbook (reference 1), we get:

Ri + Ldi/dt + 1/C (integral sign) i * dt = V(t)

$ \alpha \beta \gamma \delta \epsilon \varepsilon \zeta \eta \theta \vartheta \iota \kappa \lambda \mu \nu \xi \xi \pi \int_{}^{} \lim_{x \to \infty} \sum_{n = 1}^{\infty} \frac{\frac{d^ y}{d x^} }{}    \,dx $

References

Nilsson and Riedel, Electric Circuits 12e


\end{document}