\section{Discussion}
\label{sec:discussion}

\subsection{Summary of Problem and Results}

This paper modeled series RLC circuits using first- and second-order ODEs,
deriving analytic solutions for RC, RL, and RLC configurations.
The resistor was identified as the sole energy-dissipating element, and its
value shown to directly determine the discriminant of the characteristic
equation~\eqref{eq:char}, thereby controlling whether the circuit is
overdamped, critically damped, or underdamped.
The numerical extension then showed that stress-dependent components can
change the transient path even when the nominal values remain fixed.

\subsection{Interpretation in Terms of the Research Question}

Without resistance ($R = 0$), the characteristic roots are purely imaginary
($r = \pm j\omega_0$), yielding perpetual, undamped oscillations in an ideal
LC tank circuit.
Any nonzero $R$ introduces the decay envelope $e^{-\alpha t}$ that eventually
damps oscillations to zero.
In engineering applications: bandpass filters rely on controlled underdamping
to achieve frequency selectivity when the output is taken across the resistor,
while the capacitor-voltage response studied here behaves like a low-pass
response with possible resonant overshoot. Overdamped circuits ensure no
ringing in power supplies and ADC front-ends; critically damped designs appear
in servo systems where settling speed is paramount. A passive RLC circuit can
store and exchange energy, but a sustained oscillator also requires an active
energy source or feedback element to replace resistive losses.
This analytical framework directly addresses the research question: resistors
are essential for stabilizing oscillatory behavior and enabling practical
circuit design.
The non-ideal simulations add an important refinement: damping is not always
a constant design parameter during a transient.
When current is large, resistor heating can increase dissipation and suppress
overshoot, while inductor saturation can temporarily raise the natural
frequency and shift the timing of voltage peaks.
The resonant AC case also shows that even with resistance present, a matched
driving frequency can produce capacitor voltages much larger than the source
amplitude.

\subsection{Limitations and Future Work}

The lumped-element and linearity assumptions restrict the model's validity to
low frequencies and small signals.
Future work should incorporate:
(i) calibration of $k$ and $I_{\mathrm{sat}}$ against measured component data;
(ii) a temperature state equation for resistor self-heating rather than the
instantaneous approximation $R(i)=R_0+ki^2$;
(iii) distributed-element analysis for RF applications above $1\,\mathrm{MHz}$;
(iv) experimental validation using an oscilloscope to compare measured
$\vC(t)$ waveforms against the analytic solutions derived here; and
(v) extension to parallel RLC and more complex multi-mesh circuits modeled
by systems of ODEs.
